### 2015-09-13 doc/reports/sandia_final_impact.tex
adding reports directory in docs and adding work on final sandia fellowship report for amsi work

- Description of next steps
\label{future-work}
Currently AMSI supports only static parallel execution space discretization with non-overlapping process sets. The restriction on non-overlapping process sets is currently being worked on to allow temporary scale-reasignment while a scale is blocked waiting on an associate scale and has no work which can be conducted. However, more dynamic management of the execution space repressents the most significant current goal of the AMSI project. Implementation of this feature -- which will allow scale-tasks to grow and shrink as computational demands associated with the scale change -- is dependent on much if not all of the preceeding functionality (especially load balancing and migration features, as all relevant scale data from a process to be dynamically reassigned scales must be recovered and redistributed throughout the related scale).

Scale-balancing is the balancing of scale-tasks against eachother to minimize rendezvous time and maximize time spent in useful computation. Scale-balancing is a difficult prospect as it involves managing many dynamically-changing scales which themselves are being load-balanced using scale-sensitive load-balancing techinques, and represents a sigificant and promising area of future work. Once the ability to dynamically manage the process set associated with a scale is developed, algorithms must be developed to make the best use of this feature. Developing these algorithms repressents a major - and daunting - area of future work, as there are numerous constraints which must be taken into account in order to optimize such a scheme.

A second micro-scale simulation, where the collagen fibers are embedded in tissue is currently implemented as a standalone simulation. Work to incorporate this into the biotissue example simulation will be conducted when the correct scale-coupling terms have been developed (inclusion in the AMSI system is already accomplished, but the up- and down-scaling operations have not been developed yet so this RVE is not used). \cite{} \cite{} \cite{}

An article focusing on the implementation and use of the scale-sensitive load-balancing operations in AMSI for the biotissue example problem is being finalized and will be seeking publication soon.

- Summary of new proposals derived from work
- Follow-on funding

At present there are no current proposals associated with this work for ongoing funding, simply because funding for associated projects (Biotissue and other multi-scale projects) has already been secured. Thus work on AMSI will continue during the work on projects supported by AMSI. Several interesting multi-scale simulations have been suggested, however, particularly a combination of a Monte-Carlo code and a Mangetostatics code by a fellow researcher. At present these are in the preliminary stage of development.

- Impact on programs