### 2016-10-20 doc/articles/amsi_biotissue/draft2.tex
missing file

\title{The Adaptive Multi-scale Simulation Infrastructure}
\author{W.R. Tobin, V.W.L. Chan, M.S. Shephard}
\date{\today}
\documentclass[11pt]{article}
\bibliographystyle{siam}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[outdir=./]{epstopdf}
\usepackage{listings}
\usepackage{url}
\lstset{basicstyle=\small,
        stringstyle=\ttfamily,
        frame=top, frame=bottom,
        captionpos=t}
\begin{document}
\maketitle
\begin{abstract}
\end{abstract}
\section{Introduction}\label{introduction}

\section{Adaptive Multi-scale Simulation Infrastructure}\label{amsi}

\section{Methods}\label{methods}

\subsection{Parallel Operations}\label{parallel}

\section{Results and Discussion}\label{results}

\section{Conclusions and Future Work}\label{future_work}

\section{Acknowledgements}

\bibliography{multiscale_theory,multiscale_frameworks,load_balancing,parallel,software,biotissue}

\end{document}